---
title: "Tutorial: SLURM job management and easy parallelization from R using rSubmmiter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial: SLURM job management and easy parallelization from R using rSubmmiter}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Quick start guide 

rSubmmiter is a package that allows simple communication between R and a SLURM cluster to achieve three main tasks:

1. Easy submission, monitoring and management of individual cluster jobs.
2. Easy and fast submission of many cluster jobs, by implementing SLURM arrays.
3. Seamlessly parallelization of lapply. 

This section demonstrates the simplest way to implement these tasks

### Single job sumbssion

The creation, management and submission of individual jobs is achieved by the use of the `Job` class.
The first step is to create and object with the `Job` class by using the `new` method. On its simplest form the creation of the a `Job` object looks like this.

```{r eval=FALSE}
library("rSubmmiter")
myJob <- Job$new(commandVector = c("echo hola world!", "sleep 30"))
```

The first and only required parameter is a character vector, where each element is an independet shell command.

Then the job can be submitted to the SLURM cluster using the `$submit()` method of the `Job` object. And the `$wait()` method
can be used to wait until the job is completed of failed.

```{r eval=FALSE}
myJob$submit()
myJob$wait()
 2018-04-23 18:22:35 --- Cluster Status |  PENDING = 1 |
 2018-04-23 18:22:40 --- Cluster Status |  RUNNING = 1 |
 2018-04-23 18:23:30 --- Cluster Status |  COMPLETED = 1 |
```
